Incentivizing Exploration by Heterogeneous Users

### 2018-06-05 algorithm_regret.tex
fixed upper bound

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimates are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we define (with foresight) a phase-dependent
cutoff
$g(s) = \sqrt{\frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}}$,
and consider a regret exceeding $g(s)$ large.
\pfcomment{We use $g$ in 3 places: here, when talking about adding new linear
features in section 2, and in Lemma 11.  Maybe it's worth using different notation.}

For most of the proof, we will focus on the case $g(s) \leq \TieCutoff$.
The remaining phases are the ones with
$s \leq \frac{128}{1.8} \cdot \log(T) \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2$.
Each such phase lasts for at most $\ARMNUM \cdot \log(s)$ steps
in expectation by Lemma~\ref{lem:phase-length},
and each step incurs regret at most \MAXR.
The total expected regret incurred in these phases is at most
$100 \ARMNUM \MAXR \log^2(T) \cdot \Diam^3 d^3 \sigma^3/\TieCutoff^3$.
\pfcomment{I get the $NR$, but where does $100 \log^2(T)  D^3 d^3 \sigma^3/\hat{z}^3$ come from?  I assume this an identity for $\sum_{s=1}^m \log(s)$ I don't know --- could we add the identity we are using?}

For the remainder of the proof, we consider phases $s$
with $g(s) \leq \TieCutoff$.
We first consider agents \AgV incurring positive regret less than $g(s)$.
The second-best arm \Second{\AgV} of such an agent \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq g(s)$.
By Assumption~\ref{A1}, and because $g(s) \leq \TieCutoff$,
%Lemma~\ref{lem:sdelta},
the total measure of such agents
is $\AlmostTied{g(s)} \leq \TieDensity \cdot g(s)$,
and each such agent incurs regret at most $g(s)$.
Summing over all time steps,
and using Lemma~\ref{lem:phase-length} to bound the number of time
steps in phase $s$, 
\pfcomment{We'll have to fix a probability bug here.}
the expected total regret from such agents is at most
\begin{align}
\Expect{\sum_{s=1}^{T} \sum_{t=t_{s-1}+1}^{t_{s}} \TieDensity \cdot g^2(s)}
& \leq \sum_{s=1}^{T} \TieDensity \cdot g^2(s) \ARMNUM \log(s).
\label{equ:small_regret_bound}
\end{align}

We next bound the regret incurred in time steps with large regret.
Let the random variable $t_s$ be the final time step in phase $s$,
for each $s$.
\dkcomment{We have already used $t_s$ before, e.g., in the preceding
  formula. Do we need to redefine it here?}
\pfcomment{Above we define $t_s$ to be the start time of phase $s$.}
\dkcomment{So the new $t_s$ is the old $t_{s+1}-1$? That's a really
  confusing subtle notation change. Is there a way we can avoid this
  and just decide on one or the other, and use it consistently?}
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s}-t_{s-1}$.
By Lemma~\ref{lem:right-choice},
under \AccEU{\tau_s^k}{\frac{g(s)}{2\Diam d}}, no agent at time $t$ will incur
regret more than $g(s)$.
\dkcomment{Instead of the extra case distinction, I had thought of
  defining $g(s)$ to be the minimum of the current definition and
  \TieCutoff. Unfortunately, I think that we cannot necessarily apply
  Lemma~\ref{lem:right-choice} with that definition. Just putting that
  here in case you are thinking of simplifying the proof this way.}

To bound the probability of \AccEU{\tau_s^k}{\frac{g(s)}{2\Diam d}},
we use Lemma~\ref{lem:round-prob}
(with $x_s = x'_s = \frac{g(s)}{2 \Diam d}$
and a stopping time of $\tau_s^k$).
We first verify that this choice of $x_s, x'_s$
satisfies the assumption of Lemma~\ref{lem:round-prob} for all
$s \geq 2$,
i.e., that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x_n^2}{16 \sigma^2}}
\leq \frac{n x_n}{2 \sigma}$
for all $n \geq 2$.
Substituting that
$x_n = \frac{g(n)}{2 \Diam d} = \sqrt{\frac{32\log(T) \cdot \sigma^2}{1.8 n}}$,
the left-hand side is

@@

\begin{align*}
\sum_{s = 1}^T \sum_{t=t_s}^{t_{s+1}-1}
  \Prob{\mbox{incur regret exceeding } g(s) \mbox{ at step } t} \cdot \MAXR
& \leq
 \sum_{s = 1}^T \frac{24 \ARMNUM d}{T^2} \cdot \MAXR \cdot \ARMNUM \log(s)
\\ & \leq 12 \ARMNUM^2 d \MAXR.
\end{align*}

Adding all \dkedit{four} types of regret terms,
and substituting that 
$g^2(s) = \frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most

\begin{align*}
\lefteqn{O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
\dkedit{+ \ARMNUM \MAXR \log^2(T)}
+ \ARMNUM^2 \MAXR d
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} g^2(s) \log(s) \right)}
\\ & = 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right).\QED
\end{align*}
\end{emptyextraproof}

>>>>

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimates are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we define (with foresight) a phase-dependent
cutoff
$\EarlyCut{s} = \sqrt{\frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}}$,
and consider a regret exceeding \EarlyCut{s} large.

For most of the proof, we will focus on the case $\EarlyCut{s} \leq \TieCutoff$.
The remaining phases are the ones with
$s \leq \frac{128}{1.8} \cdot \log(T) \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2$.
Each such phase lasts for at most $\ARMNUM \cdot \log(s)$ steps
in expectation by Lemma~\ref{lem:phase-length},
and each step incurs regret at most \MAXR.
\dkedit{The total expected regret incurred in these phases is at most
$O(\ARMNUM \MAXR \cdot \log(T) \Diam^2 d^2 \sigma^2/\TieCutoff^2
\cdot \log (\log(T) \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2)
\leq O(\ARMNUM \MAXR \log^2(T) \cdot \Diam^3 d^3 \sigma^3/\TieCutoff^3)$,
by crudely bounding $\log y \leq O(\sqrt{y})$.}

For the remainder of the proof, we consider phases $s$
with $\EarlyCut{s} \leq \TieCutoff$.
We first consider agents \AgV incurring positive regret less than \EarlyCut{s}.
The second-best arm \Second{\AgV} of such an agent \AgV must satisfy
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) \leq \EarlyCut{s}$.
By Assumption~\ref{A1}, and because $\EarlyCut{s} \leq \TieCutoff$,
%Lemma~\ref{lem:sdelta},
the total measure of such agents
is $\AlmostTied{\EarlyCut{s}} \leq \TieDensity \cdot \EarlyCut{s}$,
and each such agent incurs regret at most \EarlyCut{s}.
Summing over all time steps,
and using Lemma~\ref{lem:phase-length} to bound the number of time
steps in phase $s$, 
\pfcomment{We'll have to fix a probability bug here.}
the expected total regret from such agents is at most
\begin{align}
\Expect{\sum_{s=1}^{T} \sum_{t=t_{s-1}+1}^{t_{s}} \TieDensity \cdot \EarlyCut[2]{s}}
& \leq \sum_{s=1}^{T} \TieDensity \cdot g^2(s) \ARMNUM \log(s).
\label{equ:small_regret_bound}
\end{align}

We next bound the regret incurred in time steps with large regret.
Let the random variable $t_s$ be the final time step in phase $s$,
for each $s$.
\dkcomment{We have already used $t_s$ before, e.g., in the preceding
  formula. Do we need to redefine it here?}
\pfcomment{Above we define $t_s$ to be the start time of phase $s$.}
\dkcomment{So the new $t_s$ is the old $t_{s+1}-1$? That's a really
  confusing subtle notation change. Is there a way we can avoid this
  and just decide on one or the other, and use it consistently?}
Define the stopping times $\tau_{s}^{k}$ to be the \Kth{k}
time step in the \Kth{s} phase,
with $\tau_{s}^{k} = \infty$ when phase $s$ has fewer than $k$ steps,
i.e., $k > t_{s}-t_{s-1}$.
By Lemma~\ref{lem:right-choice},
under \AccEU{\tau_s^k}{\frac{\dkedit{\EarlyCut{s}}}{2\Diam d}}, no agent at time $t$ will incur
regret more than \EarlyCut{s}.
\dkcomment{Instead of the extra case distinction, I had thought of
  defining \EarlyCut{s} to be the minimum of the current definition and
  \TieCutoff. Unfortunately, I think that we cannot necessarily apply
  Lemma~\ref{lem:right-choice} with that definition. Just putting that
  here in case you are thinking of simplifying the proof this way.}

To bound the probability of \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}},
we use Lemma~\ref{lem:round-prob}
(with $x_s = x'_s = \frac{\EarlyCut{s}}{2 \Diam d}$
and a stopping time of $\tau_s^k$).
We first verify that this choice of $x_s, x'_s$
satisfies the assumption of Lemma~\ref{lem:round-prob} for all
$s \geq 2$,
i.e., that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x_n^2}{16 \sigma^2}}
\leq \frac{n x_n}{2 \sigma}$
for all $n \geq 2$.
Substituting that
$x_n = \frac{\EarlyCut{n}}{2 \Diam d} = \sqrt{\frac{32\log(T) \cdot \sigma^2}{1.8 n}}$,
the left-hand side is

@@

\begin{align*}
\sum_{s = 1}^T \sum_{t=t_s}^{t_{s+1}-1}
  \Prob{\mbox{incur regret exceeding } \EarlyCut{s} \mbox{ at step } t} \cdot \MAXR
& \leq
 \sum_{s = 1}^T \frac{24 \ARMNUM d}{T^2} \cdot \MAXR \cdot \ARMNUM \log(s)
\\ & \leq 12 \ARMNUM^2 d \MAXR.
\end{align*}

Adding all \dkedit{four} types of regret terms,
and substituting that 
$\EarlyCut[2]{s} = \frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most

\begin{align*}
\lefteqn{O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
\dkedit{+ \ARMNUM \MAXR \log^2(T)}
+ \ARMNUM^2 \MAXR d
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)}
\\ & = 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right).\QED
\end{align*}
\end{emptyextraproof}

### 2018-06-05 problem_setting.tex
resolving some easy edits
fixed upper bound

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
Although the model is linear, it permits additional flexibility
through the addition of derived attributes that are nonlinear
transformation of some original attribute vector.
One can simply define $\Ag{j} = g_j(\AgV)$ for some known
nonlinear functions $g_j$ and $j=d+1,\ldots,d'$, and then increase the
number of attributes to $d'$.

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).
Notice that based on Assumption~\ref{A1} below,
\dkedit{\Best{\AgV} is unique with probability 1.}

>>>>

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
Although the model is linear, it permits additional flexibility
through the addition of derived attributes that are nonlinear
transformation of some original attribute vector.
One can simply define \dkedit{$\Ag{j} = h_j(\AgV)$ for some known
nonlinear functions $h_j$} and $j=d+1,\ldots,d'$, and then increase the
number of attributes to $d'$.

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).
Notice that based on Assumption~\ref{A1} below,
\Best{\AgV} is unique with probability 1.
